---
title: Analysing spatial data using R
subtitle: Brunei R User Group Meetup 🇧🇳
date: '9 March 2023'
author:
  - name: Haziq Jamil
    orcid: 0000-0003-3298-1010
    email: haziq.jamil@ubd.edu.bn
    affiliations:
      - name: Universiti Brunei Darussalam
# website:
#   repo-url: https://github.com/bruneir/br-spatial
---

```{r}
#| label: setup
#| include: false
library(tidyverse)
library(bruneimap)
theme_set(theme_bw())
```

# Preliminaries

- Brunei R User Group

# Introduction

- Why spatial analysis?
- What is spatial data? Points, lines, polygons.
- How can R help?

# Getting started with R

- RStudio overview
- Importing data
- Installing packages
- The tidyverse
- The [`{bruneimap}`](https://github.com/propertypricebn/bruneimap) package

```{r}
ggplot(kpg_sf) +
  geom_sf(aes(fill = mukim), col = "gray50") +
  geom_sf(data = mkm_sf, col = "black", lwd = 0.5, fill = NA) +
  geom_sf(data = filter(kpg_sf, is.na(mukim)), fill = "gray70", col = "gray70") +
  scale_fill_viridis_d(option = "turbo") +
  theme(legend.position = "none") 
```


# Brunei population

::: {.callout-tip title="What we'll learn"}

- Represent statistical data using colour mapping symbology (choropleth)
- Use `ggplot2::geom_label()` or `ggrepel::geom_label_repel()` to add labels to the map
- Using a binned colour scale, e.g. `ggplot2::geom_scale_fill_viridis_b()` 

:::


```{r}
bn_pop_sf <- 
  left_join(
    kpg_sf, 
    bn_census2021, 
    by = join_by(id, kampong, mukim, district
  ))

kpg_labels_sf <-
  bn_pop_sf |>
  arrange(desc(population)) |>
  slice_head(n = 10)

bn_pop_sf |>
  # filter(population > 50) |>
  ggplot() +
  geom_sf(aes(fill = population), col = NA, alpha = 0.8) +
  geom_sf(data = kpg_sf, fill = NA, col = "black") +
  ggrepel::geom_label_repel(
    data = kpg_labels_sf,
    aes(label = kampong, geometry = geometry),
    stat = "sf_coordinates",
    inherit.aes = FALSE,
    box.padding = 1,
    size = 2,
    max.overlaps = Inf
  ) +
  scale_fill_viridis_b(
    name = "Population",
    na.value = NA,
    labels = scales::comma,
    breaks = c(0, 50, 100, 1000, 5000, 10000, 15000)
    # limits = c(0, 12000)
  ) +
  theme_bw()
```

